---
title: "Assignment STAT702"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r global-options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE,
                      fig.width=6, fig.height=4)
```

```{r setup}
# Load libraries
library(tidyverse)
library(lubridate)

# Read in data sets and convert to tibbles
reviews_data <- read.csv("reviews_data.csv")
reviews_data <- as_tibble(reviews_data)

sales_data <- read.csv("sales_data.csv")
sales_data <- as_tibble(sales_data)
```

**Group 5**: Hitarth Asrani and Genevieve Connell

**Product name**: BIC Round Stic Xtra Life Ballpoint Pen, Medium Point
(1.0mm), Red, 12-Count

**Sales sku_id**: 219884 

**Reviews asin**: B00006IE7J


# 1 Analysis of Sales Data

## 1(a) For the product (sku_id) which has been assigned to your group (see page 6), compute the total monthly sales from January 2011 -- July 2013. Present your results in an appropriate plot and write 2 -- 3 sentences describing your results.

Hint: This will require some "wrangling" of the variable week. To do this, format week as a date and then use the appropriate lubridate function to extract the month.

Marking Criteria

-   Total monthly sales have been correctly computed and are displayed
    in an appropriate plot.

-   Description of results/plot is correct and provides useful insights.

-   Plot is constructed using ggplot2 and has appropriate titles,
    labels, scales etc.**


## Answer

From Jan 2011 - July 2013 `r sum(monthly_sales$total_units_sold)` units of sku 219844 were sold with a mean monthly sale of `r mean(monthly_sales$total_units_sold`. Over this period there was a significant outlier with `r outlier$total_units_sold` units sold on `r outlier$date`. 

50% of monthly sales lie between `r summary(monthly_sales$total_units_sold)[2]` and `r summary(monthly_sales$total_units_sold)[5]`, this is the interquartile range. 

Monthly sales are plotted below, no trend or seasonal pattern is evident in this plot. 

```{r 1a}
# Create a summary table, grouped by month and year (single column with m/y)
sales_data %>% 
  filter(sku_id == 219844) %>%
  mutate(week = as_date(week, format = "%d/%m/%y"),
         date = format(week, "%m/%y"),
         date = my(date)) %>%
  group_by(date) %>%
  summarise(total_units_sold = sum(units_sold))-> monthly_sales

# Line plot of monthly sales
ggplot(monthly_sales) +
  geom_line(aes(x = date, y = total_units_sold)) +
  xlab("Date") +
  ylab("Units Sold") +
  ggtitle("Monthly sales for product 219844 (Jan 2011 - July 2013)")

# Compute basic stats
summary(monthly_sales$total_units_sold)

# Identify outlier
monthly_sales %>% 
  arrange(desc(total_units_sold)) %>% 
  slice_max(total_units_sold) -> outlier
```


## 1(b) The GM Sales wants to know which stores are performing well and which are not, in terms of product sales. For the product (sku_id) which has been assigned to your group, use appropriate summary statistics and plots to investigate sales performance across the stores and write 2 -- 3 paragraphs summarising your findings.

Hint: You will need to decide what it means for a store to be
"performing well" and how you will evaluate this using the data.

Marking criteria

-   Sales performance is clearly defined.

-   Written summary includes relevant and appropriate summary statistics
    and plots.

-   Plot/s are constructed using ggplot2 and have appropriate titles,
    labels, scales etc.

-   Descriptions of results and plots are correct and provides useful
    insights.
    
## Answer

Quarterly sales have been calculated for each store. The mean number of units sold by a store in a quarter is `r mean(quarterly_sales)` with an interquartile range of `r q1` - `r q3`. This means that 50% of store quarterly sales lie within the range of `r q1` - `r q3` and this range is used as a benchmark for comparing store performance. 

If a store sells more than `r q3` units then their quarterly performance is classified as 'High'. If a store sells less than `r q1` then their quarterly performance is classified as 'Low'. If they sell within the interquartile range they are classified as 'Average'. 

In 2013 obesrvations only include January - July. This means there is only partial data for quarter 3 and quarter 4 so these quarters are exculded from the following analysis. 

```{r 1b}
# Summarise overall sales per store
sales_data %>% 
  filter(sku_id == 219844) %>%
  mutate(week = as_date(week, format = "%d/%m/%y"),
         date = format(week, "%m/%y"),
         date = my(date),
         year = year(date),
         quarter = quarter(date),
         store_id = as.character(store_id)) %>% 
  filter(between(date, as_date("2011-01-01"), as_date("2013-06-30"))) %>% 
  group_by(store_id) %>%
  summarise(store_total = sum(units_sold))-> store_sales

# Rank stores based on total sales
store_sales %>% 
  arrange(desc(store_total)) %>% 
  mutate(ranking = seq(1:28)) -> store_sales
```


```{r}
# Monthly sales per store
sales_data %>% 
  filter(sku_id == 219844) %>%
  mutate(week = as_date(week, format = "%d/%m/%y"),
         date = format(week, "%m/%y"),
         date = my(date),
         store_id = as.character(store_id)) %>%
  group_by(date, store_id) %>%
  summarise(total_units_sold = sum(units_sold))-> monthly_sales

# Classify monthly performance
monthly_sales %>% 
  mutate(performance = ifelse
         (total_units_sold < summary(monthly_sales$total_units_sold)[2], "Low", 
                              (ifelse
                               (total_units_sold > summary(monthly_sales$total_units_sold)[5],
                                 "High", 
                                 "Average")))) -> monthly_sales

# categorise month based on if product is on display or not
sales_data %>% 
  filter(sku_id == 219844) %>%
  mutate(week = as_date(week, format = "%d/%m/%y"),
         date = format(week, "%m/%y"),
         date = my(date),
         year = year(date),
         quarter = quarter(date),
         store_id = as.character(store_id),
         category = ifelse(is_featured_sku == "1", "Featured", 
                              (ifelse(is_display_sku == "1", "Display", "None")))) -> sales_2

# Add overall store ranking to monthly sales
monthly_sales %>% 
  left_join(store_sales, by = "store_id") %>% 
  select(store_id, date, total_units_sold, performance, ranking) -> monthly_sales



``` 

```{r}

``` 
  
```{r}  
# Summarise by quarterly sales for each store
sales_data %>% 
  filter(sku_id == 219844) %>%
  mutate(week = as_date(week, format = "%d/%m/%y"),
         date = format(week, "%m/%y"),
         date = my(date),
         year = year(date),
         quarter = quarter(date),
         store_id = as.character(store_id)) %>% 
  filter(between(date, as_date("2011-01-01"), as_date("2013-06-30"))) %>% 
  unite("quarter", year, quarter, remove = FALSE, sep = "/") %>%
  group_by(quarter, store_id) %>% 
  summarise(total_units_sold = sum(units_sold))  -> quarterly_sales

# Calculate quarterly quartiles
q1 <- summary(quarterly_sales$total_units_sold)[2]
q3 <- summary(quarterly_sales$total_units_sold)[5]

# Add performance classification
quarterly_sales %>% 
  mutate(performance = ifelse(total_units_sold < q1, "Low", 
                              (ifelse(total_units_sold > q3, "High", "Average")))) -> quarterly_sales

# Add ranking to quarterly sales
quarterly_sales %>% 
  left_join(store_sales, by = "store_id") %>% 
  select(quarter, total_units_sold, store_id, performance, ranking) -> quarterly_sales

# All plots
quarterly_sales %>% 
  ggplot()+
  geom_point(aes(x = quarter, y = total_units_sold, colour = performance)) + 
  facet_wrap(~ store_id, ncol = 6) +
  scale_colour_manual(values = c("#80CAFF", "#4BC979", "#E76F51")) +
  scale_x_discrete(labels = c("2011", "", "", "", "2012", "", "", "", "2013", ""))

```

```{r}
# Plot bottom third
quarterly_sales %>% 
  filter(ranking >18) %>% 
  ggplot()+
  geom_point(aes(x = quarter, y = total_units_sold, colour = performance)) + 
  facet_wrap(~ store_id, ncol = 3) +
  scale_colour_manual(values = c("#80CAFF", "#E76F51")) +
  scale_x_discrete(labels = c("2011", "", "", "", "2012", "", "", "", "2013", ""))

# Plot middle third
quarterly_sales %>% 
  filter(ranking <18, ranking >9) %>% 
  ggplot()+
  geom_point(aes(x = quarter, y = total_units_sold, colour = performance)) + 
  facet_wrap(~ store_id, ncol = 3) +
  scale_colour_manual(values = c("#80CAFF", "#4BC979", "#E76F51")) +
  scale_x_discrete(labels = c("2011", "", "", "", "2012", "", "", "", "2013", ""))

# Plot rank top  third
quarterly_sales %>% 
  filter(ranking < 9) %>% 
  ggplot()+
  geom_point(aes(x = quarter, y = total_units_sold, colour = performance)) + 
  facet_wrap(~ store_id, ncol = 3) +
  scale_colour_manual(values = c("#80CAFF", "#4BC979", "#E76F51")) +
  scale_x_discrete(labels = c("2011", "", "", "", "2012", "", "", "", "2013", ""))



```

```{r}
# categorise based on if product is on display or not
sales_data %>% 
  filter(sku_id == 219844) %>%
  mutate(week = as_date(week, format = "%d/%m/%y"),
         date = format(week, "%m/%y"),
         date = my(date),
         year = year(date),
         quarter = quarter(date),
         store_id = as.character(store_id),
         category = ifelse(is_featured_sku == "1", "Featured", 
                              (ifelse(is_display_sku == "1", "Display", "None")))) -> sales_2

# Add ranking 
sales_2 %>% 
  left_join(store_sales, by = "store_id") -> sales_2
  
sales_2 %>% 
  ggplot()+
  geom_point(aes(x = week, y = units_sold, shape = category)) + 
  facet_wrap(~ store_id, ncol = 6)


sales_2 %>% 
  filter(ranking == 1) %>% 
  ggplot()+
  geom_point(aes(x = week, y = units_sold, colour = category)) + 
  facet_wrap(~ store_id, ncol = 6)

sales_2 %>% 
  geom_point(aes(x= week, y = units_sold, colour = category))


# bottom third
sales %>% 
  filter(ranking >18) %>% 
  ggplot()+
  geom_point(aes(x = week, y = total_units_sold, shape = category)) + 
  facet_wrap(~ store_id, ncol = 3) +
  scale_colour_manual(values = c("#80CAFF", "#E76F51")) +
  scale_x_discrete(labels = c("2011", "", "", "", "2012", "", "", "", "2013", ""))

```

# Question 2

## (a) The Operations Manager is interested in studying an EOQ model for
    product 216233, based on sales in 2012. The setup and holding costs
    are known to be 130 per order and 1.50 per unit per year,
    respectively.


## i)  Determine the best order quantity in such a way that the costs are minimised. Write 1 -- 2 paragraphs summarising your findings.

Marking criteria 
• Number of orders during a year, number of days
between orders, and the total annual inventory cost are correctly computed and included in the findings. 
• The paragraphs clearly explain your findings. 
• Assumptions of the EOQ model are clearly stated

```{r 2ai}
# Annual demand in 2012 for product 216233
sales_data %>% 
   mutate(week = as_date(week, format = "%d/%m/%y"),
         year = year(week)) %>% 
  filter(sku_id == "216233", year == 2012) %>%
  select(units_sold) %>% 
  sum()-> A 

k <- 130 # set up costs per order
h <- 1.5 # holding costs per unit

Q <- sqrt(2*k*A / h) # optimum order quantity 
Q <- round(Q, 0)


t <- sqrt(2*k/(A*h)) # inventory cycle 
t <- round(365*t, 0)

Tc <- k*A/Q + h*Q/2 # annual inventory cost


plot(0, xlim = c(0, 2*t), ylim = c(0, Q), type = "n",
     xlab = "Time (days)", ylab = "Stock level",
     main = "Inventory cycles for 216233")
segments(x0 = c(0,t), y0 = Q, x1 = c(t,2*t), y1 = 0) # diagonals
segments(x0 = c(0,t), y0 = 0, x1 = c(0,t), y1 = Q)  # vertical


```

Optimum order quantity is `r Q`

Inventory cycle is `r t`

Annual inventory cost is `r Tc`


Assumptions of EOQ model:

Known and constant demand

No lead time - orders arrive instantaneously

No back orders

## ii) The Operations Manager is also interested in studying a model in
    which backorders are permitted. According to its estimates, the cost
    of backorders is approximately 5% of the total price (price per
    unit). Determine the best order quantity in the sense that inventory
    costs are minimised. Write 1 -- 2 paragraphs summarising your
    findings and plot the first two inventory cycles.

Marking criteria 
• The optimum order quantity, maximum level of stock, optimum time between orders, proportion of time the company have to take backorders, and total annual inventory cost are correctly computed and included in your answer. 
• The paragraphs clearly explain your findings.
• Assumptions of the model are clearly stated. 
• The first two inventory cycles are correctly plotted

```{r 2aii}
# Annual demand in 2012 for product 216233
sales_data %>% 
   mutate(week = as_date(week, format = "%d/%m/%y"),
         year = year(week)) %>% 
  filter(sku_id == "216233", year == 2012) %>%
  select(units_sold) %>% 
  sum()-> A 

k <- 130 # set up costs per order
h <- 1.5 # holding costs per unit

sales_data %>% 
   mutate(week = as_date(week, format = "%d/%m/%y"),
         year = year(week)) %>% 
  filter(sku_id == "216233", year == 2012) %>% 
  select(total_price)-> x # create vector of prices for product 216233

p <- mean(x$total_price)*0.05 # back order cost per unit (0.05 of total price)

Q <- sqrt(2*k*A/h) * sqrt((p+h)/p)
Q <- round(Q, 0) # Optimum ordering quantity

S <- sqrt(2*k*A/h) * sqrt(p/(p+h))
S <- round(S, 0) # Optimum maximum inventory level

t <- round(365*Q/A, 0) # Optimum time between orders

t1 <- round(365*S/A, 0) # Time supplier is in stock
pt1 <- 100*(t-t1)/t # Proportion of time supplier taking back orders

Tc <- k*A/Q + h/2 * S^2/Q + p/2 * (Q-S)^2 /Q # Total inventory cost

bo <- Q*(pt1/100) # back orders before reordering

# Plotting
plot(0, xlim = c(0, 2*t), ylim = c(-bo, Q), type = "n",
     xlab = "Time (days)", ylab = "Stock level",
     main = "Inventory cycles for 216233")
segments(x0 = c(0,t), y0 = Q, x1 = c(t,2*t), y1 = -bo) # diagonals
segments(x0 = c(0,t), y0 = -bo, x1 = c(0,t), y1 = Q)  # vertical
segments(x0 = c(0,t), y0 = 0, x1 = c(t,2*t), y1 = 0) # horizontal
```

Optimum ordering quantity `r Q`

Optimum maximum inventory level `r S`

Optimum time between orders `r t`

Proportion of time taking back orders `r pt1`

Total inventory cost `r Tc`


Assumptions of model:

known and constant demand

no lead time

orders arrive instantaneously

back orders allowed

demand can be backordered when no stock


## iii) Plot the inventory cycles associated with the model in part ii and compare with the observed inventory levels in 2012, assuming actual demand during 2012, and the order frequency and order quantity from the model. Write 2 – 3 sentences describing your plot.

Marking criteria
• The inventory levels from the model and data are correctly plotted.
• Accurate and insightful comments are made about the plot.
• Note: This is a bonus question. The maximum mark that could be awarded for this
project is 100

```{r 2aiii}
# Calculate weekly sales in 2012 for product 216233
sales_data %>% 
   mutate(week = as_date(week, format = "%d/%m/%y"),
         year = year(week))%>% 
  filter(sku_id == "216233", year == 2012) %>%
  group_by(week) %>% 
  summarise(quantity = sum(units_sold)) -> weekly_sales

# create data frame with reorder quantities and dates
seq(from = 0, to = 365, by = t) %>% 
  as_tibble() %>% 
  mutate(week = as_date("2012-01-01") + value, quantity = Q) %>% 
  select(week, quantity) -> inventory

# combine data frames and create a running total
weekly_sales %>% 
  mutate(quantity = quantity * (-1)) %>% 
  rbind(inventory) %>% 
  arrange(week) %>% 
  mutate(inventory = cumsum(quantity), 
         type = (ifelse(quantity > 0, "inventory", "sales"))) -> inventory2

ggplot(inventory2) +
  geom_line(aes(x = week, y = inventory)) +
  geom_hline(yintercept = 0)

```

The Operations Manager is considering the option of a multi-period inventory model. The company, as a policy, is not willing to tolerate more than 5% chance of a stock-out. The Operations Manager has estimated that the annual holding cost is 6.50 per unit and the ordering cost is 20.50 per order.

i. Calculate a multi-period inventory model for product 216425, based on the 2012 sales data. Create plot/s of the weekly average demand of this product. Use the costs stated in part (b)above. Write a paragraph explaining the results of your model and the plot/s.

Hint: Use the weekly demand to estimate the demand during a one-week lead time.

Marking criteria
• The optimal order quantity, safety stock, expected annual cost, orders per years are correctly computed and included in your answer.
• The paragraph clearly explains your findings.
• The assumption of normality for the demand during a one-week lead time is discussed.
• The weekly average demand of this product is correctly plotted and discussed

```{r 2bi}
# Calculate weekly sales in 2012 for product 216425
sales_data %>% 
   mutate(week = as_date(week, format = "%d/%m/%y"),
         year = year(week))%>% 
  filter(sku_id == "216425", year == 2012) %>%
  group_by(week) %>% 
  summarise(quantity = sum(units_sold)) -> weekly_sales

mean <- mean(weekly_sales$quantity)
sd <- sd(weekly_sales$quantity)

# Expected annual demand 
D <- mean*52


# subset total_price to calculate average total_price
sales_data %>% 
   mutate(week = as_date(week, format = "%d/%m/%y"),
         year = year(week)) %>% 
  filter(sku_id == "216425", year == 2012) %>% 
  select(total_price)-> y # create vector of prices for product 216425

# purchase price
pp <- mean(y$total_price)

# ordering cost
co <- 20.5 # per order

# annual holding cost
ch <- 6.5 # per unit

# optimal ordering quantity
Q = round(sqrt(2 * D * co / ch), 0)

# time between orders 
t <- round(365*Q/D, 0)

# orders per year
n <- round(D/Q, 0)

#  Reorder point r that allows 5% chance of a stock-out
r <- qnorm(0.95, mean, sd)

x <- seq(from= 0,to=4000,len=100)
plot(x,dnorm(x,mean=mean,sd=sd ),type="l",lwd=3,col="black") +
abline(v = r)

# safety stock
Qs <- r - mean

## Expected annual cost

# Holding cost (normal stock)
hc_normal <- Q * ch / 2

# Holding cost (safety stock)
hc_safety <- Qs * ch

# Ordering cost
oc <- D * co / Q

# Total annual cost 
tac <- hc_normal + hc_safety + oc

# annual cost with no safety stock
tac2 <- hc_normal + oc
```

Demand during a one week lead time has been estimated based off 2012 weekly demand. The mean weekly demand from 2012 and standard deviation have been applied to a normal distribution to estimate demand for this multi-inventory model. 

As the plot below shows, the actual demand for 2012 does not perfectly follow a normal distribution. It is difficult to determine the distribution of weekly demand based on only one year of observations.It is recommended that more data is used to for a more accurate model of demand distribution. 

Based on 2012, where mean weekly demand was `r mean` with standard deviation of `r sd`, the expected annual demand is estimated to be `r D`. 

Given this annual demand and the costs of holding and reordering stock, the recommended multi-inventory model is to order `r Q` units whenever the order quantity reaches the reorder  point of `r r` units. Approximately `r n` orders will be placed per year. Safety stock is `r Q`. 

This approach ensures roughly 95% of the time the `r r` units will be able to satisfy demand during the lead time. 

The expected annual costs are `r tac` per year. If demand was certain the annual costs would only be `r tac`, the additional cost of holding safety stock is the cost of uncertain demand. 

```{r 2bi plot}
# Plot observations against estimated lead-time distribution

# set bin width and number of observations for plotting
bw <- 300
n_obs = sum(!is.na(weekly_sales$quantity))

g <- ggplot(weekly_sales, aes(quantity))  + 
geom_histogram(aes(y = ..density..), binwidth = bw, colour = "black") + 
stat_function(fun = dnorm, args = list(mean = mean(weekly_sales$quantity), sd = sd(weekly_sales$quantity)))

ybreaks = seq(0,20,1) 

g + scale_y_continuous("Density", sec.axis = sec_axis(
  trans = ~ . * bw * n_obs, name = "Counts", breaks = ybreaks))

```

## 2.b.ii. Investigate the use of a multi-period inventory model for the product which has been assigned to your group, based on the 2012 sales data. Create plot/s of the weekly average demand of this product. Use the costs stated in part (b) above.

## Discuss the assumptions of the model and suggest a solution, in case of finding any problems. Write a paragraph explaining the results of your findings and the plot.

```{r 2bii}
# Calculate weekly sales in 2012 for product 219844
sales_data %>% 
   mutate(week = as_date(week, format = "%d/%m/%y"),
         year = year(week))%>% 
  filter(sku_id == "219844", year == 2012) %>%
  group_by(week) %>% 
  summarise(quantity = sum(units_sold)) -> weekly_sales

mean <- mean(weekly_sales$quantity)
sd <- sd(weekly_sales$quantity)

# Expected annual demand 
D <- mean*52


# subset total_price to calculate average total_price
sales_data %>% 
   mutate(week = as_date(week, format = "%d/%m/%y"),
         year = year(week)) %>% 
  filter(sku_id == "219844", year == 2012) %>% 
  select(total_price)-> y # create vector of prices for product 219844

# purchase price
pp <- mean(y$total_price)

# ordering cost
co <- 20.5 # per order

# annual holding cost
ch <- 6.5 # per unit

# optimal ordering quantity
Q = round(sqrt(2 * D * co / ch), 0)

# time between orders 
t <- round(365*Q/D, 0)

# orders per year
n <- round(D/Q, 0)

#  Reorder point r that allows 5% chance of a stock-out
r <- qnorm(0.95, mean, sd)

x <- seq(from=0,to=2000,len=100)
plot(x,dnorm(x,mean=mean,sd=sd ),type="l",lwd=3,col="black") +
abline(v = r)

# safety stock
Qs <- r - mean

## Expected annual cost

# Holding cost (normal stock)
hc_normal <- Q * ch / 2

# Holding cost (safety stock)
hc_safety <- Qs * ch

# Ordering cost
oc <- D * co / Q

# Total annual cost 
tac <- hc_normal + hc_safety + oc

# annual cost with no safety stock
tac2 <- hc_normal + oc

```

Demand during a one week lead time has been estimated based off 2012 weekly demand. The mean weekly demand from 2012 and standard deviation have been applied to a normal distribution to estimate demand for this multi-inventory model. 

As the plot below shows the observed demand does not follow normal distribution with mean `r mean` `r sd`. It is recommended that a more suitable model is used to estimate deand. 

Based on the flawed normal model, where mean weekly demand was `r mean` with standard deviation of `r sd`, the expected annual demand is estimated to be `r D`. 

Given this annual demand and the costs of holding and reordering stock, the recommended multi-inventory model is to order `r Q` units whenever the order quantity reaches the reorder  point of `r r` units. Approximately `r n` orders will be placed per year. Safety stock is `r Q`. 

This approach ensures roughly 95% of the time the `r r` units will be able to satisfy demand during the lead time. 

The expected annual costs are `r tac` per year. If demand was certain the annual costs would only be `r tac`, the additional cost of holding safety stock is the cost of uncertain demand. 

```{r 2bii plot}
# Plot observations against estimated lead-time distribution

# set bin width and number of observations for plotting
bw <- 300
n_obs = sum(!is.na(weekly_sales$quantity))

g <- ggplot(weekly_sales, aes(quantity))  + 
geom_histogram(aes(y = ..density..), binwidth = bw, colour = "black") + 
stat_function(fun = dnorm, args = list(mean = mean(weekly_sales$quantity), sd = sd(weekly_sales$quantity)))

ybreaks = seq(0,42,2) 

g + scale_y_continuous("Density", sec.axis = sec_axis(
  trans = ~ . * bw * n_obs, name = "Counts", breaks = ybreaks))


```


# Question 3

## Question 3a

```{r 3a}
reviews_data %>%
  filter(asin == "B00006IE7J") -> my_reviews

head(my_reviews)
summary(my_reviews$overall)


# Stats - plots 
ggplot(data = my_reviews) +
  geom_bar(mapping = aes(x=overall, fill=overall))


```
