---
title: "Assignment STAT702"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Product name: BIC Round Stic Xtra Life Ballpoint Pen, Medium Point (1.0mm), Red, 12-Count
Sales sku_id: 219844
Reviews asin: B00006IE7J 


#1 Analysis of Sales Data

## 1(a) For the product (sku_id) which has been assigned to your group (see page 6), compute the total monthly sales from January 2011 – September 2013. Present your results in an appropriate plot and write 2 – 3 sentences describing your results.

```{r load_libs}
library(tidyverse)
library(lubridate)
library(dplyr)
library(ggplot2)
```

1 Read data

```{r read data}
reviews_data = read.csv("reviews_data.csv") %>% as_tibble()
sales_data = read.csv("sales_data.csv") %>% as_tibble()

head(reviews_data)
head(sales_data)
```


```{r filter_data}

sales_data %>%
  filter(sku_id == 219844) %>%
  mutate(week = as_date(week, format = "%d/%m/%y"),
         date = my(format(week, "%m/%y"))) %>% 
  group_by(date) %>% 
  summarise(total_units = sum(units_sold)) %>% 
  filter(between(date, as_date("2011-01-01"), as_date("2013-09-30"))) -> sales_jan11_to_sep13
sales_jan11_to_sep13
```
```{r plot-sales}
ggplot(sales_jan11_to_sep13) +
  geom_point(mapping = aes(x = date, y = total_units)) +
   geom_line(mapping = aes(x = date, y = total_units)) +
  ggtitle("Monthly Sales (Jan 11 - Sep 13) for SKU 219844")
```

In the above plot we can see that the sales of the product #219844 decreases sharply during after January and July. We also notice a sharp increase in sales around the months of December and June. This trend continues yearly.

## 1(b)

## Question 3a

```{r filter_reviews}
my_reviews = reviews_data %>% filter(asin == "B00006IE7J")

reviewRange = 1:5
head(my_reviews)
summary(my_reviews$overall)
# Stats - plots 
ggplot(data = my_reviews) +
  geom_bar(mapping = aes(x=overall, fill=overall))
```

## Question 3b
```{r sentiment_data}
library(tidytext)

sentiment_dat = my_reviews %>% select(document.id, overall, verified, reviewText, summary)
head(sentiment_dat)

# Tidying
sentiment_dat %>% 
  unnest_tokens(output = reviewWord, input = reviewText) -> tidy_sentiment

tidy_sentiment %>% 
  unnest_tokens(output = summaryWord, input = summary) %>% print(n = 5) -> tidy_sentiment

# Add Row Number
tidy_sentiment %>% mutate(linenumber = row_number()) -> tidy_sentiment
head(tidy_sentiment)

# Remove Stop Words
data("stop_words")
tidy_sentiment %>% anti_join(stop_words)

```



