---
title: "Assignment STAT702"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1 Analysis of Sales Data

##(a) For the product (sku_id) which has been assigned to your group (see page 6), compute the total monthly sales from January 2011 – September 2013. Present your results in an appropriate plot and write 2 – 3 sentences describing your results.

Hint: This will require some “wrangling” of the variable week. To do this, format week as a date and then use the appropriate lubridate function to extract the month.

Marking Criteria

 *  Total monthly sales have been correctly computed and are displayed in an appropriate plot.
 
 *  Description of results/plot is correct and provides useful insights.
 
 *  Plot is constructed using ggplot2 and has appropriate titles, labels, scales etc.
 
```{r}
# Load libraries
library(tidyverse)
library(lubridate)

```

```{r}
# Read in data and convert to tibbles
reviews_data <- read.csv("reviews_data.csv")
reviews_data <- as_tibble(reviews_data)

sales_data <- read.csv("sales_data.csv")
sales_data <- as_tibble(sales_data)
```

```{r}
# Subset data lines with sku_id == 219844

sales_219844 %>% 
        filter(sku_id == 219844) %>% 
        mutate(date = as_date(week, format = "%d/%m/%y"))-> sales_data_2


``` 
